Analysis

### 2024-12-16 production/prod_analysis-fix-t5.Rmd
change to quarto + add ola 7

# Df variables nivel 1 _________________________________________________________
df_study1 <- df_study_long_t6_comunas %>% dplyr::select(ola,
                                               just_educ, des_perc, merit_talent, merit_effort,
                                               educ, quintil1:sexo)

df_study1$ola <- factor(df_study1$ola,
         labels = c("2016","2017","2018","2019", "2022"))
df_study1$ola <- 
sjlabelled::set_label(x = df_study1$ola,
                      label = "Wave")

@@

lab_study1 <- sjlabelled::get_label(df_study1) 
df_study1 <- df_study1 %>% na.omit() %>% sjlabelled::set_label(lab_study1)

df<- dfSummary(x = dplyr::select(df_study1, merit_talent, merit_effort, des_perc, educ, quintil1, ess, edad, sexo, pos_id),
               plain.ascii = FALSE,
               style = "grid",
               tmp.img.dir = "/tmp",
               graph.magnif = TRUE,
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = F,    # missing
               graph.col = T, # plot
               valid.col = F, # n valido
               col.widths = c(700,10,10,10)
               )
df$Variable <- NULL # delete variable column
summarytools::view(df,
     file = here::here("output/tables/desc02.html")
     ) # Ver tabla en un archivo HTML
webshot::webshot(url =here::here("output/tables/desc02.html"),
        file =here::here("output/tables/desc02.png")) # Remover archivo HTML
# print(df, method = "render")
```

```{r}
load(file = here::here("input/data-proc/df_study_t6_ind_comunas.RData"))
df_study1 <- df_study_long_t6_comunas %>% dplyr::select(idencuesta, ola, ponderador_long_total, segmento, estrato,
                                               just_educ, des_perc, merit_effort, merit_talent,
                                               educ, quintil1:sexo) %>% na.omit()
df_study1 <- df_study1[complete.cases(df_study1$segmento), ]

@@

```{r alluvial}
colors<- c("#f1eef6ff","#bdc9e1ff","#b3b3b3ff","#74a9cfff","#0570b0ff")
alluvial_justeduc <- ggplot(datos.justeduc, aes(x = ola, fill = just_educ, stratum = just_educ,
                             alluvium = idencuesta, y = porcentaje))+
    ggalluvial::geom_flow(alpha = .66) + 
    ggalluvial::geom_stratum(linetype = 0) +
    scale_y_continuous(labels = scales::percent) + 
    ylab(label = NULL) +
    xlab(label = NULL) + 
    theme(#legend.position = 'top',
          legend.title = element_blank(),
          plot.caption = element_text(hjust = 1)) +
    scale_fill_manual(values = colors) +
    geom_shadowtext(data = etiquetas.justeduc, 
              aes(label = ifelse(porcentaje > 0 , scales::percent(porcentaje, accuracy = .1),"")),
              position = position_stack(vjust = .5),
              show.legend = FALSE,
              size = 4,
              color = rep('white'),
              bg.colour='grey30')+
  scale_x_discrete(labels = c(2016, 2017, 2018, 2019, 2022)) +
  labs(caption = "Source: own elaboration with pooled data from ELSOC 2016-2022 (n = 8042)")

@@

```{r bivariate-plots}
df_study1 <- df_study_long_t6_comunas %>% dplyr::select(idencuesta, ola, ponderador_long_total, segmento, estrato,
                                               just_educ, des_perc, merit_effort, merit_talent,
                                               educ, quintil1:sexo) %>% na.omit()
merit_plot <- df_study1 %>% 
  select(merit_effort, merit_talent, des_perc, just_educ) %>% 
  sjPlot::plot_stackfrq(weight.by = df_study1$ponderador_long_total,
                        expand.grid =TRUE,
                      geom.colors = "PuBu",
                      show.total = FALSE, 
                      vjust=rep(c("bottom", "top", "bottom", "top", "bottom"),4),
                      legend.labels = c("Strongly disagree", "Disagree", "Neither agree
nor disagree", "Agree", "Strongly agree"),
                      axis.labels = c("Merit effort", "Merit talent", "Inequality 
perception", "Justification 
of 
educational inequality")
                      ) +
  theme_light(base_size = 16) +
  theme(legend.position="bottom", text = element_text(size = 14),
        plot.caption = element_text(hjust = 1)) +
  labs(caption = "Source: own elaboration with pooled data from ELSOC 2016-2022 (n = 8042)")
merit_plot

@@

```{r}

range <- c(1, 1)

p_just_educ<-
  dplyr::select(df_study1,
                just_educ) %>%
  sjPlot::plot_likert(geom.colors = "PuBu",
                      geom.size = 0.8,
                      catcount = 5,
                      grid.range  =  range,
                      show.prc.sign = T,
                      reverse.colors = F,
                      reverse.scale = T,
                      show.n = FALSE,
                      intercept.line.color = "white")  +
  ggplot2::theme(legend.position = "bottom",
    text = ggplot2::element_text(size = 16))+
  ggplot2::scale_fill_manual(
    values = colors,
    breaks = c(1,2,3,4,5),
  ) +
  theme(legend.position = "none")

p_des_perc<-
  dplyr::select(df_study1,
                des_perc) %>%
  sjPlot::plot_likert(geom.colors = "PuBu",
                      geom.size = 0.8,
                      catcount = 5,
                      grid.range  =  range,
                      show.prc.sign = T,
                      reverse.colors = F,
                      reverse.scale = T,
                      show.n = FALSE,
                      intercept.line.color = "white")  +
  ggplot2::theme(legend.position = "bottom",
    text = ggplot2::element_text(size = 16))+
  ggplot2::scale_fill_manual(
    values = colors,
    breaks = c(1,2,3,4,5),
  ) +
  theme(legend.position = "none")

p_effort<-
  dplyr::select(df_study1,
                merit_effort) %>%
  sjPlot::plot_likert(geom.colors = "PuBu",
                      geom.size = 0.8,
                      catcount = 5,
                      grid.range  =  range,
                      show.prc.sign = T,
                      reverse.colors = T,
                      reverse.scale = T,
                      show.n = FALSE,
                      intercept.line.color = "white")  +
  ggplot2::theme(legend.position = "bottom",
    text = ggplot2::element_text(size = 16))+
  ggplot2::scale_fill_manual(
    values = colors,
    breaks = c(1,2,3,4,5),
  ) +
  theme(legend.position = "none")

p_talent<-
  dplyr::select(df_study1,
                merit_talent) %>%
  sjPlot::plot_likert(geom.colors = "PuBu",
                      geom.size = 0.8,
                      catcount = 5,
                      grid.range  =  range,
                      show.prc.sign = T,
                      reverse.colors = F,
                      reverse.scale = T,
                      show.n = FALSE, 
                      intercept.line.color = "white")  +
  ggplot2::theme(legend.position = "bottom",
    text = ggplot2::element_text(size = 16))+
  ggplot2::scale_fill_manual(
    values = colors,
    breaks = c(1,2,3,4,5), 
    labels = c(
      "Strongly disagree",
      "Disagree",
      "Neither disagree nor agree",
      "Agree",
      "Strongly agree"
    )
  )
```

##### Create plot
```{r}
pacman::p_load(cowplot)
t <- 0.1
s <- 0.45
m <- 0.4
xs <- 0.25

caption <- 
  stringr::str_wrap(
    paste0("Note: Own elaboration based on ELSOC Survey ",
           "(n=",dim(na.omit(df_study1))[1],"). "),
    width = 100)

title <- cowplot::ggdraw() + 
  cowplot::draw_label(
    "Frequencies",
    fontface = 'bold',
    size = 15,
    x = 0,
    hjust = 0
  )

likert1<-   
  cowplot::add_sub(
    cowplot::plot_grid(p_just_educ,
                       p_des_perc,
                       p_effort,
                       p_talent,
                       align = "hv",
                       axis = "a",
                       nrow = 7, #6 plots
                       ncol = 1, #1 column   
                       rel_heights = c(
                                       s,s,s,s,
                                       s,
                                       s) #fig size within plot
    ),
    label = "", # Caption: source and sample size
    size = 11,  # Caption font size
    hjust = 0# Caption align 
  ) %>% 
  cowplot::ggdraw()
jpeg(here::here("output/graphs/figure1.jpg"),
    # units = "cm",
    # res = 100,
    width = 800,
    height = 800)
likert1
dev.off()
```

justification education vs year

```{r}
just_educ_year = df_study1 %>%
  group_by(ola) %>%
  summarise(education = mean(just_educ, na.rm = T))

categorias_x <- c(2016, 2017, 2018, 2019, 2022)

educ_year_plot <- just_educ_year_survey %>% ggplot(aes(ola, education)) +
  theme_bw(base_size = 14) +
  geom_point(size = 4)+
  ylab("Justification of inequality") +
  xlab("Wave") +
  scale_x_discrete(labels = categorias_x) +
  ylim(1,3)
educ_year_plot

ggsave(educ_year_plot, file="output/graphs/educ_year_plot.png", width = 7, height = 7)
```

@@

```{r}
load(file = here::here("input/data-proc/df_study_t6_ind_comunas.RData"))
df_study1 <- df_study_long_t6_comunas %>% dplyr::select(idencuesta, ola, ponderador_long_total, segmento, estrato,
                                               just_educ, des_perc, merit_effort, merit_talent,
                                               educ, quintil1:sexo) %>% na.omit()
df_study1 <- df_study1[complete.cases(df_study1$segmento), ]

@@

categorias_x <- c(2016, 2017, 2018, 2019, 2022)
vars_years$variable <- factor(vars_years$variable, levels = c("education", "inequality", "talent", "effort"), labels=  c("Justification
educational
inequality
(dependent)", "Inequality
perception", "Merit: Talent", "Merit: Effort"),
ordered = TRUE)

@@

## Crear el gráfico principal
years_plot <- ggplot(vars_years, aes(ola, value, color = variable, shape = variable, linetype = variable)) +
  theme_bw(base_size = 14) +
  geom_point(size = 4) +
  geom_line(aes(group = variable)) +
  geom_errorbar(aes(ymin = value_low, ymax = value_upp), width = 0.2) +
  ylab("Mean") +
  xlab("Wave") +
  ylim(1, 5) +
  scale_x_discrete(labels = categorias_x) +
  scale_color_manual(labels = categorias_z, name = "Variable", values =  c("#bdc9e1ff","#0570b0ff","#74a9cfff","#b3b3b3ff")) +
  scale_shape_manual(labels = categorias_z, name = "Variable", values = c("square", "circle", "triangle", "diamond")) +
  scale_linetype_manual(labels = categorias_z, name = "Variable", 
                        values = c("solid", "dashed", "dotted", "dotdash")) +  # Definir diferentes tipos de línea
   theme(legend.position = "none") +
  coord_cartesian(xlim = c(1, 6)) +  # Ampliar el eje x
  scale_x_discrete(labels = c(2016, 2017, 2018, 2019, 2022, ""))  # Personalizar etiquetas del eje x

@@

# Añadir etiquetas personalizadas
years_plot <- years_plot +
  geom_point(data = label_data, aes(label_x - 0.05, label_y, color = variable, shape = variable), size = 4) +  # Capa de puntos para los símbolos de forma
  geom_text(data = label_data, aes(label_x, label_y, label = variable), color = "black", hjust = 0, nudge_x = 0.05, nudge_y = 0, size = 4)  +
  labs(caption = "Source: own elaboration with pooled data from ELSOC 2016-2022 (n = 8042)")

@@

ggsave("output/graphs/years_plot.png", plot = years_plot, height = 6, width = 7, dpi = 300)

@@

```{r}
# Calcular cuantiles para las variables

boxplot <- df_study_long_t6_comunas %>% select(just_educ, des_perc, merit_talent, merit_effort, ola) %>% as.data.frame()

boxplot <- boxplot %>% 
  pivot_longer(cols = c("just_educ", 
                        "des_perc", 
                        "merit_talent", 
                        "merit_effort"), 
               names_to = "variable",
               values_to = "value") %>% na.omit()

boxplot$variable <- factor(boxplot$variable,
                   levels = c("just_educ", 
                        "des_perc", 
                        "merit_talent", 
                        "merit_effort"),
                   labels = c("Justification
educational
inequality", "Inequality
perception", "Merit: Talent", "Merit: Effort"))

boxplot_plot <- boxplot %>%
  ggplot(aes(x = as.factor(ola), y = value, fill=variable, color=variable)) + 
  geom_boxplot() +
  scale_y_continuous(name = "Value", limits = c(1, 5)) +
  scale_x_discrete(name = "Wave", labels=categorias_x) +
  labs(fill = "Variable", color = "Variable") +
  theme_bw() +
  theme(axis.text.x = element_text(margin = margin(b = 15)))
boxplot_plot

ggsave(boxplot_plot, file="output/graphs/boxplot_plot.png", height = 7, width = 7)
```

```{r cor-plot, fig.dim=c(10,6)}
rm(list=ls())
if (!require("pacman")) install.packages("pacman") # instalar pacman
                            # cargar librerias
pacman::p_load(dplyr,       # Manipulacion de datos
               knitr,
               kableExtra,
               sjPlot,
               sjmisc,
               summarytools,
               lavaan
               )
load(file = here::here("input/data-proc/df_study_t6_ind_comunas.RData"))
df_study1 <- df_study_long_t6_comunas %>% dplyr::select(idencuesta, ola, ponderador_long_total, segmento, estrato,
                                               just_educ, des_perc, merit_effort, merit_talent,
                                               educ, quintil1:sexo) %>% na.omit()
# Full sample_______________________________________________________________________
mat_full<- 
df_study1 %>%
  # filter(ola==1) %>%
  select(just_educ, # distributive justice
         des_perc,
         merit_effort, merit_talent # meritocracy
         ) %>% 
  psych::mixedCor()

cor_full <- as.matrix(mat_full$rho)
diag(cor_full) = NA #set diagonal values to NA
# Set Row names of the matrix
rownames(cor_full) <- c("A. Justification of educational inequality",
                        "B. Inequality perception",
                     "C. People are rewarded for their efforts",
                     "D. People are rewarded for their intelligence")
#set Column names of the matrix
colnames(cor_full) <-c("(A)", "(B)","(C)","(D)")
#Plot the matrix using corrplot
png("output/graphs/corr.png",width=600,height=300, pointsize = 10)
corrplot::corrplot(cor_full,
  method = "color",
  addCoef.col = "#000390",
  type = "upper",
  tl.col = "black",
  col=colorRampPalette(c("white","#0068DC"))(12),
  bg = "white",
  na.label = "-")
# Agregar caption
grid::grid.text("Source: own elaboration with pooled data 
from ELSOC 2016-2022 (n = 8042)", x = 0.01, y = 0.05, just = "left", gp = grid::gpar(fontsize = 10, col = "black"))
dev.off()
# Wave 1_______________________________________________________________________
mat_1<- 
df_study1_long_comunas %>%
  filter(ola==1) %>%
  select(just_educ, # distributive justice
         merit_effort, merit_talent, # meritocracy
         ahead_family, ahead_ambition, ahead_educ, ahead_hardwork # go ahead
         ) %>% 
  psych::mixedCor()

cor_1 <- as.matrix(mat_1$rho)
diag(cor_1) = NA #set diagonal values to NA
# Set Row names of the matrix
rownames(cor_1) <- c("A. Education distributive justice",
                     "B. People are rewarded for their efforts",
                     "C. People are rewarded for their intelligence",
                     "D. Go ahead: wealthy family",
                     "E. Go ahead: Having ambition",
                     "F. Go ahead: Good level of education",
                     "G. Go ahead: Hard work")
#set Column names of the matrix
colnames(cor_1) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)")
#Plot the matrix using corrplot
corrplot::corrplot(cor_1,
  method = "color",
  addCoef.col = "#000390",
  type = "upper",
  tl.col = "black",
  col=colorRampPalette(c("white","#0068DC"))(12),
  bg = "white",
  na.label = "-",
  title = "\n Wave 1")

# Wave 2_______________________________________________________________________
mat_2<-   
df_study1_long_comunas %>%
  filter(ola==2) %>% 
  select(just_educ, # distributive justice
         merit_effort, merit_talent, # meritocracy
         ahead_family, ahead_ambition, ahead_educ, ahead_hardwork # go ahead
         ) %>% 
  psych::mixedCor()

cor_2 <- as.matrix(mat_2$rho)
diag(cor_2) = NA #set diagonal values to NA
# Set Row names of the matrix
rownames(cor_2) <- c("A. Education distributive justice",
                     "B. People are rewarded for their efforts",
                     "C. People are rewarded for their intelligence",
                     "D. Go ahead: wealthy family",
                     "E. Go ahead: Having ambition",
                     "F. Go ahead: Good level of education",
                     "G. Go ahead: Hard work")
#set Column names of the matrix
colnames(cor_2) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)")
#Plot the matrix using corrplot
corrplot::corrplot(cor_2,
  method = "color",
  addCoef.col = "#000390",
  type = "upper",
  tl.col = "black",
  col=colorRampPalette(c("white","#0068DC"))(12),
  bg = "white",
  na.label = "-",
  title = "\n Wave 2")

# Wave 3_______________________________________________________________________
mat_3 <- 
df_study1_long_comunas %>%
  filter(ola==3) %>% 
  select(just_educ, # distributive justice
         merit_effort, merit_talent, # meritocracy
         ahead_family, ahead_ambition, ahead_educ, ahead_hardwork # go ahead
         ) %>% 
  psych::mixedCor()

cor_3 <- as.matrix(mat_3$rho)
diag(cor_3) = NA #set diagonal values to NA
# Set Row names of the matrix
rownames(cor_3) <- c("A. Education distributive justice",
                     "B. People are rewarded for their efforts",
                     "C. People are rewarded for their intelligence",
                     "D. Go ahead: wealthy family",
                     "E. Go ahead: Having ambition",
                     "F. Go ahead: Good level of education",
                     "G. Go ahead: Hard work")
#set Column names of the matrix
colnames(cor_3) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)")
#Plot the matrix using corrplot
corrplot::corrplot(cor_3,
  method = "color",
  addCoef.col = "#000390",
  type = "upper",
  tl.col = "black",
  col=colorRampPalette(c("white","#0068DC"))(12),
  bg = "white",
  na.label = "-",
  title = "\n Wave 3")

# Wave 4_______________________________________________________________________
mat_4 <- 
df_study1_long_comunas %>%
  filter(ola==4) %>% 
  select(just_educ, # distributive justice
         merit_effort, merit_talent, # meritocracy
         ahead_family, ahead_ambition, ahead_educ, ahead_hardwork # go ahead
         ) %>% 
  psych::mixedCor()

cor_4 <- as.matrix(mat_4$rho)
diag(cor_4) = NA #set diagonal values to NA
# Set Row names of the matrix
rownames(cor_4) <- c("A. Education distributive justice",
                     "B. People are rewarded for their efforts",
                     "C. People are rewarded for their intelligence",
                     "D. Go ahead: wealthy family",
                     "E. Go ahead: Having ambition",
                     "F. Go ahead: Good level of education",
                     "G. Go ahead: Hard work")
#set Column names of the matrix
colnames(cor_4) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)")
#Plot the matrix using corrplot
corrplot::corrplot(cor_4,
  method = "color",
  addCoef.col = "#000390",
  type = "upper",
  tl.col = "black",
  col=colorRampPalette(c("white","#0068DC"))(12),
  bg = "white",
  na.label = "-",
  title = "\n Wave 4")

# Wave 6_______________________________________________________________________
mat_6 <- 
df_study1_long_comunas %>%
  filter(ola==6) %>% 
  select(just_educ, # distributive justice
         merit_effort, merit_talent, # meritocracy
         ahead_family, ahead_ambition, ahead_educ, ahead_hardwork # go ahead
         ) %>% 
  psych::mixedCor()

cor_6 <- as.matrix(mat_6$rho)
diag(cor_6) = NA #set diagonal values to NA
# Set Row names of the matrix
rownames(cor_6) <- c("A. Education distributive justice",
                     "B. People are rewarded for their efforts",
                     "C. People are rewarded for their intelligence",
                     "D. Go ahead: wealthy family",
                     "E. Go ahead: Having ambition",
                     "F. Go ahead: Good level of education",
                     "G. Go ahead: Hard work")
#set Column names of the matrix
colnames(cor_6) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)")
#Plot the matrix using corrplot
corrplot::corrplot(cor_6,
  method = "color",
  addCoef.col = "#000390",
  type = "upper",
  tl.col = "black",
  col=colorRampPalette(c("white","#0068DC"))(12),
  bg = "white",
  na.label = "-",
  title = "\n Wave 6")
```

>>>>

# Df variables nivel 1 _________________________________________________________
df_study1 <- df_study_long_t7_comunas %>% dplyr::select(ola,
                                               just_educ, des_perc, merit_talent, merit_effort,
                                               educ, quintil1:sexo)

df_study1$ola <- factor(df_study1$ola,
         labels = c("2016","2017","2018","2019", "2022", "2023"))
df_study1$ola <- 
sjlabelled::set_label(x = df_study1$ola,
                      label = "Wave")

@@

lab_study1 <- sjlabelled::get_label(df_study1) 
df_study1 <- df_study1 %>% sjlabelled::set_label(lab_study1)

df<- dfSummary(x = dplyr::select(df_study1, ola, merit_talent, merit_effort, des_perc, educ, quintil1, ess, edad, sexo, pos_id),
               plain.ascii = FALSE,
               style = "grid",
               tmp.img.dir = "/tmp",
               graph.magnif = TRUE,
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = F,    # missing
               graph.col = T, # plot
               valid.col = F, # n valido
               col.widths = c(700,10,10,10)
               )
df$Variable <- NULL # delete variable column
summarytools::view(df,
     file = here::here("output/tables/desc02.html")
     ) # Ver tabla en un archivo HTML
webshot::webshot(url =here::here("output/tables/desc02.html"),
        file =here::here("output/tables/desc02.png")) # Remover archivo HTML
# print(df, method = "render")
```

```{r}
load(file = here::here("input/data-proc/df_study_t6_ind_comunas.RData"))
df_study1 <- df_study_long_t7_comunas %>% dplyr::select(idencuesta, ola, ponderador_long_total, segmento, estrato,
                                               just_educ, des_perc, merit_effort, merit_talent,
                                               educ, quintil1:sexo) %>% na.omit()
df_study1 <- df_study1[complete.cases(df_study1$segmento), ]

@@

```{r alluvial}
colors<- c("#f1eef6ff","#bdc9e1ff","#b3b3b3ff","#74a9cfff","#0570b0ff")
alluvial_justeduc <- ggplot(datos.justeduc, aes(x = ola, fill = just_educ, stratum = just_educ,
                             alluvium = idencuesta, y = porcentaje))+
    ggalluvial::geom_flow(alpha = .66) + 
    ggalluvial::geom_stratum(linetype = 0) +
    scale_y_continuous(labels = scales::percent) + 
    ylab(label = NULL) +
    xlab(label = NULL) + 
    theme(#legend.position = 'top',
          legend.title = element_blank(),
          plot.caption = element_text(hjust = 1)) +
    scale_fill_manual(values = colors) +
    geom_shadowtext(data = etiquetas.justeduc, 
              aes(label = ifelse(porcentaje > 0 , scales::percent(porcentaje, accuracy = .1),"")),
              position = position_stack(vjust = .5),
              show.legend = FALSE,
              size = 4,
              color = rep('white'),
              bg.colour='grey30')+
  scale_x_discrete(labels = c(2016, 2017, 2018, 2019, 2022, 2023)) +
  labs(caption = "Source: own elaboration with pooled data from ELSOC 2016-2023 (n = 9519)")

@@

```{r bivariate-plots}
df_study1 <- df_study_long_t7_comunas %>% dplyr::select(idencuesta, ola, ponderador_long_total, segmento, estrato,
                                               just_educ, des_perc, merit_effort, merit_talent,
                                               educ, quintil1:sexo) %>% na.omit()
merit_plot <- df_study1 %>% 
  select(merit_effort, merit_talent, des_perc, just_educ) %>% 
  sjPlot::plot_stackfrq(weight.by = df_study1$ponderador_long_total,
                        expand.grid =TRUE,
                      geom.colors = "PuBu",
                      show.total = FALSE, 
                      vjust=rep(c("bottom", "top", "bottom", "top", "bottom"),4),
                      legend.labels = c("Strongly disagree", "Disagree", "Neither agree
nor disagree", "Agree", "Strongly agree"),
                      axis.labels = c("Merit effort", "Merit talent", "Inequality 
perception", "Justification 
of 
educational inequality")
                      ) +
  theme_light(base_size = 16) +
  theme(legend.position="bottom", text = element_text(size = 14),
        plot.caption = element_text(hjust = 1)) +
  labs(caption = "Source: own elaboration with pooled data from ELSOC 2016-2022 (n = 9519)")
merit_plot

@@

@@

```{r}
load(file = here::here("input/data-proc/df_study_t6_ind_comunas.RData"))
df_study1 <- df_study_long_t7_comunas %>% dplyr::select(idencuesta, ola, ponderador_long_total, segmento, estrato,
                                               just_educ, des_perc, merit_effort, merit_talent,
                                               educ, quintil1:sexo) %>% na.omit()
df_study1 <- df_study1[complete.cases(df_study1$segmento), ]

@@

categorias_x <- c(2016, 2017, 2018, 2019, 2022, 2023)
vars_years$variable <- factor(vars_years$variable, levels = c("education", "inequality", "talent", "effort"), labels=  c("Justification
educational
inequality
(dependent)", "Inequality
perception", "Merit: Talent", "Merit: Effort"),
ordered = TRUE)

@@

## Crear el gráfico principal
years_plot <- ggplot(vars_years, aes(ola, value, color = variable, shape = variable, linetype = variable)) +
  theme_bw(base_size = 14) +
  geom_point(size = 4) +
  geom_line(aes(group = variable)) +
  geom_errorbar(aes(ymin = value_low, ymax = value_upp), width = 0.2) +
  ylab("Mean") +
  xlab("Wave") +
  ylim(1, 5) +
  scale_x_discrete(labels = categorias_x) +
  scale_color_manual(labels = categorias_z, name = "Variable", values =  c("#bdc9e1ff","#0570b0ff","#74a9cfff","#b3b3b3ff")) +
  scale_shape_manual(labels = categorias_z, name = "Variable", values = c("square", "circle", "triangle", "diamond")) +
  scale_linetype_manual(labels = categorias_z, name = "Variable", 
                        values = c("solid", "dashed", "dotted", "dotdash")) +  # Definir diferentes tipos de línea
   theme(legend.position = "none") +
  coord_cartesian(xlim = c(1, 7)) +  # Ampliar el eje x
  scale_x_discrete(labels = c(2016, 2017, 2018, 2019, 2022, 2023, ""))  # Personalizar etiquetas del eje x

@@

# Añadir etiquetas personalizadas
years_plot <- years_plot +
  geom_point(data = label_data, aes(label_x + 1, label_y, color = variable, shape = variable), size = 4) +  # Capa de puntos para los símbolos de forma
  geom_text(data = label_data, aes(label_x, label_y, label = variable), color = "black", hjust = 0, nudge_x = 1.1, nudge_y = 0, size = 4)  +
  labs(caption = "Source: own elaboration with pooled data from ELSOC 2016-2023 (n = 9519)")

@@

ggsave("output/graphs/years_plot.png", plot = years_plot, height = 6, width = 7, dpi = 300)
```

```{r cor-plot, fig.dim=c(10,6)}
rm(list=ls())
if (!require("pacman")) install.packages("pacman") # instalar pacman
                            # cargar librerias
pacman::p_load(dplyr,       # Manipulacion de datos
               knitr,
               kableExtra,
               sjPlot,
               sjmisc,
               summarytools,
               lavaan
               )
load(file = here::here("input/data-proc/df_study_t6_ind_comunas.RData"))
df_study1 <- df_study_long_t7_comunas %>% dplyr::select(idencuesta, ola, ponderador_long_total, segmento, estrato,
                                               just_educ, des_perc, merit_effort, merit_talent,
                                               educ, quintil1:sexo) %>% na.omit()
# Full sample_______________________________________________________________________
mat_full<- 
df_study1 %>%
  # filter(ola==1) %>%
  select(just_educ, # distributive justice
         des_perc,
         merit_effort, merit_talent # meritocracy
         ) %>% 
  psych::mixedCor()

cor_full <- as.matrix(mat_full$rho)
diag(cor_full) = NA #set diagonal values to NA
# Set Row names of the matrix
rownames(cor_full) <- c("A. Justification of educational inequality",
                        "B. Inequality perception",
                     "C. People are rewarded for their efforts",
                     "D. People are rewarded for their intelligence")
#set Column names of the matrix
colnames(cor_full) <-c("(A)", "(B)","(C)","(D)")
#Plot the matrix using corrplot
png("output/graphs/corr.png",width=600,height=300, pointsize = 10)
corrplot::corrplot(cor_full,
  method = "color", # Cambia los círculos por color completo de cada cuadrante
  addCoef.col = "#000390", # Color de los coeficientes
  type = "upper", # Deja solo las correlaciones de arriba
  tl.col = "black", # COlor letras, rojo por defecto
  na.label = "-")
# Agregar caption
grid::grid.text("Source: own elaboration with pooled data 
from ELSOC 2016-2023 (n = 9519)", x = 0.01, y = 0.05, just = "left", gp = grid::gpar(fontsize = 10, col = "black"))
dev.off()
# Wave 1_______________________________________________________________________
# mat_1<- 
# df_study1_long_comunas %>%
#   filter(ola==1) %>%
#   select(just_educ, # distributive justice
#          merit_effort, merit_talent, # meritocracy
#          ahead_family, ahead_ambition, ahead_educ, ahead_hardwork # go ahead
#          ) %>% 
#   psych::mixedCor() 
# 
# cor_1 <- as.matrix(mat_1$rho)
# diag(cor_1) = NA #set diagonal values to NA
# # Set Row names of the matrix
# rownames(cor_1) <- c("A. Education distributive justice",
#                      "B. People are rewarded for their efforts",
#                      "C. People are rewarded for their intelligence",
#                      "D. Go ahead: wealthy family",
#                      "E. Go ahead: Having ambition",
#                      "F. Go ahead: Good level of education",
#                      "G. Go ahead: Hard work")
# #set Column names of the matrix
# colnames(cor_1) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)")
# #Plot the matrix using corrplot
# corrplot::corrplot(cor_1,
#   method = "color",
#   addCoef.col = "#000390",
#   type = "upper",
#   tl.col = "black",
#   col=colorRampPalette(c("white","#0068DC"))(12),
#   bg = "white",
#   na.label = "-",
#   title = "\n Wave 1")
# 
# # Wave 2_______________________________________________________________________
# mat_2<-   
# df_study1_long_comunas %>%
#   filter(ola==2) %>% 
#   select(just_educ, # distributive justice
#          merit_effort, merit_talent, # meritocracy
#          ahead_family, ahead_ambition, ahead_educ, ahead_hardwork # go ahead
#          ) %>% 
#   psych::mixedCor() 
# 
# cor_2 <- as.matrix(mat_2$rho)
# diag(cor_2) = NA #set diagonal values to NA
# # Set Row names of the matrix
# rownames(cor_2) <- c("A. Education distributive justice",
#                      "B. People are rewarded for their efforts",
#                      "C. People are rewarded for their intelligence",
#                      "D. Go ahead: wealthy family",
#                      "E. Go ahead: Having ambition",
#                      "F. Go ahead: Good level of education",
#                      "G. Go ahead: Hard work")
# #set Column names of the matrix
# colnames(cor_2) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)")
# #Plot the matrix using corrplot
# corrplot::corrplot(cor_2,
#   method = "color",
#   addCoef.col = "#000390",
#   type = "upper",
#   tl.col = "black",
#   col=colorRampPalette(c("white","#0068DC"))(12),
#   bg = "white",
#   na.label = "-",
#   title = "\n Wave 2")
# 
# 
# # Wave 3_______________________________________________________________________
# mat_3 <- 
# df_study1_long_comunas %>%
#   filter(ola==3) %>% 
#   select(just_educ, # distributive justice
#          merit_effort, merit_talent, # meritocracy
#          ahead_family, ahead_ambition, ahead_educ, ahead_hardwork # go ahead
#          ) %>% 
#   psych::mixedCor()  
# 
# cor_3 <- as.matrix(mat_3$rho)
# diag(cor_3) = NA #set diagonal values to NA
# # Set Row names of the matrix
# rownames(cor_3) <- c("A. Education distributive justice",
#                      "B. People are rewarded for their efforts",
#                      "C. People are rewarded for their intelligence",
#                      "D. Go ahead: wealthy family",
#                      "E. Go ahead: Having ambition",
#                      "F. Go ahead: Good level of education",
#                      "G. Go ahead: Hard work")
# #set Column names of the matrix
# colnames(cor_3) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)")
# #Plot the matrix using corrplot
# corrplot::corrplot(cor_3,
#   method = "color",
#   addCoef.col = "#000390",
#   type = "upper",
#   tl.col = "black",
#   col=colorRampPalette(c("white","#0068DC"))(12),
#   bg = "white",
#   na.label = "-",
#   title = "\n Wave 3")
# 
# # Wave 4_______________________________________________________________________
# mat_4 <- 
# df_study1_long_comunas %>%
#   filter(ola==4) %>% 
#   select(just_educ, # distributive justice
#          merit_effort, merit_talent, # meritocracy
#          ahead_family, ahead_ambition, ahead_educ, ahead_hardwork # go ahead
#          ) %>% 
#   psych::mixedCor() 
# 
# 
# cor_4 <- as.matrix(mat_4$rho)
# diag(cor_4) = NA #set diagonal values to NA
# # Set Row names of the matrix
# rownames(cor_4) <- c("A. Education distributive justice",
#                      "B. People are rewarded for their efforts",
#                      "C. People are rewarded for their intelligence",
#                      "D. Go ahead: wealthy family",
#                      "E. Go ahead: Having ambition",
#                      "F. Go ahead: Good level of education",
#                      "G. Go ahead: Hard work")
# #set Column names of the matrix
# colnames(cor_4) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)")
# #Plot the matrix using corrplot
# corrplot::corrplot(cor_4,
#   method = "color",
#   addCoef.col = "#000390",
#   type = "upper",
#   tl.col = "black",
#   col=colorRampPalette(c("white","#0068DC"))(12),
#   bg = "white",
#   na.label = "-",
#   title = "\n Wave 4")
# 
# 
# # Wave 6_______________________________________________________________________
# mat_6 <- 
# df_study1_long_comunas %>%
#   filter(ola==6) %>% 
#   select(just_educ, # distributive justice
#          merit_effort, merit_talent, # meritocracy
#          ahead_family, ahead_ambition, ahead_educ, ahead_hardwork # go ahead
#          ) %>% 
#   psych::mixedCor() 
# 
# cor_6 <- as.matrix(mat_6$rho)
# diag(cor_6) = NA #set diagonal values to NA
# # Set Row names of the matrix
# rownames(cor_6) <- c("A. Education distributive justice",
#                      "B. People are rewarded for their efforts",
#                      "C. People are rewarded for their intelligence",
#                      "D. Go ahead: wealthy family",
#                      "E. Go ahead: Having ambition",
#                      "F. Go ahead: Good level of education",
#                      "G. Go ahead: Hard work")
# #set Column names of the matrix
# colnames(cor_6) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)")
# #Plot the matrix using corrplot
# corrplot::corrplot(cor_6,
#   method = "color",
#   addCoef.col = "#000390",
#   type = "upper",
#   tl.col = "black",
#   col=colorRampPalette(c("white","#0068DC"))(12),
#   bg = "white",
#   na.label = "-",
#   title = "\n Wave 6")

@@
